---
title: "Survival Analysis Report"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    html_encoding: "UTF-8"
encoding: UTF-8
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(encoding = "UTF-8")

skill_dir <- "/Users/tongzj/Library/Application Support/Claude/local-agent-mode-sessions/skills-plugin/208a1ca7-880b-4ac7-a2c9-05379a00effd/b921ad37-a4f4-428a-acd2-b0faf2258a66/skills/statistical-analysis"
source(file.path(skill_dir, "scripts/two_groups_combined.R"))

if (!requireNamespace("knitr",      quietly = TRUE)) install.packages("knitr")
if (!requireNamespace("kableExtra", quietly = TRUE)) install.packages("kableExtra")
library(knitr)
library(kableExtra)
library(dplyr)
```

```{r computation, echo=FALSE}
data      <- read.csv("/Users/tongzj/Desktop/SIBDS dsci/llm project/SIBDS research project (Jenni, Lilian)/mock_datasets/whas100_12.csv")
group_1   <- data$agecat
group_2   <- data$gender
event     <- data$fstat
time      <- data$lenfol
threshold <- 0.05

res                  <- lrt_test(group_1, group_2, event, time, threshold)
lr_chisq             <- res$logrank["chisq"]
lr_p                 <- res$logrank["p.value"]
lr_df                <- 7
survival_mean_CI     <- res$survival_mean_CI
survival_quantile_CI <- res$survival_quantile_CI
group_summary        <- res[[6]]

p_label <- if (lr_p < 0.001) {
  "< 0.001 (&#42;&#42;&#42;)"
} else if (lr_p < 0.01) {
  "< 0.01 (&#42;&#42;)"
} else if (lr_p < 0.05) {
  paste0(round(lr_p, 4), " (&#42;)")
} else {
  as.character(round(lr_p, 4))
}

decision <- if (lr_p < threshold) "Reject Null Hypothesis" else "Fail to Reject Null Hypothesis"
```

## Hypothesis Test Summary

```{r hypothesis-summary, echo=FALSE, results='asis'}
cat(paste0(
  "**Null Hypothesis:** Survival curves are identical across all groups of agecat and gender.\n\n",
  "**Alternative Hypothesis:** At least one survival curve differs.\n\n",
  "| Statistic | Value |\n",
  "|---|---|\n",
  "| &chi;&sup2; statistic | ", round(lr_chisq, 2), " |\n",
  "| Degrees of freedom | ", lr_df, " |\n",
  "| p-value | ", p_label, " |\n",
  "| Decision | ", decision, " at &alpha; = ", threshold, " |\n\n"
))
```

## Background and Methods

This survival analysis examines whether age category (`agecat`) and gender jointly affect time-to-event (`lenfol`, days from hospital admission to death), with death (`fstat = 1`) as the event of interest and administrative censoring (`fstat = 0`) indicating patients alive at the end of follow-up. The analysis includes 8 strata formed by crossing age category (4 levels: [32,60), [60,70), [70,80), [80,92]) &times; gender (Female = 0, Male = 1) across 100 patients. A **log-rank test with Bonferroni correction** was selected because two categorical covariates are present, producing more than two comparison groups. The three key assumptions are: (1) **Proportional Hazards** &mdash; the hazard ratio between any two groups remains constant over time; (2) **Non-Informative Censoring** &mdash; whether an observation is censored is unrelated to the underlying risk of the event; and (3) **Independent Observations** &mdash; each subject's survival time is independent of all others.

The log-rank test yielded &chi;&sup2;(`r lr_df`) = `r round(lr_chisq, 2)`, p = `r p_label`. The decision is to **`r decision`** at significance level &alpha; = `r threshold`, indicating there is a statistically significant difference in survival across at least one of the 8 agecat &times; gender strata.

## Quantile Table

```{r quantile-table, echo=FALSE}
n_events <- group_summary %>%
  mutate(strata = paste0(group_1, ".", group_2))

qtbl <- data.frame(
  Group   = gsub("group=", "", as.character(survival_quantile_CI$group)),
  n       = n_events$n[match(
              gsub("group=", "", as.character(survival_quantile_CI$group)),
              paste0(n_events$group_1, ".", n_events$group_2))],
  Events  = n_events$event_happened[match(
              gsub("group=", "", as.character(survival_quantile_CI$group)),
              paste0(n_events$group_1, ".", n_events$group_2))],
  `Median (95% CI)` = paste0(
    ifelse(is.na(survival_quantile_CI$q50), "NR", round(survival_quantile_CI$q50, 2)),
    " (", ifelse(is.na(survival_quantile_CI$q50_lower), "NR", round(survival_quantile_CI$q50_lower, 2)),
    "&ndash;", ifelse(is.na(survival_quantile_CI$q50_upper), "NR", round(survival_quantile_CI$q50_upper, 2)), ")"),
  `Q25 (95% CI)` = paste0(
    ifelse(is.na(survival_quantile_CI$q25), "NR", round(survival_quantile_CI$q25, 2)),
    " (", ifelse(is.na(survival_quantile_CI$q25_lower), "NR", round(survival_quantile_CI$q25_lower, 2)),
    "&ndash;", ifelse(is.na(survival_quantile_CI$q25_upper), "NR", round(survival_quantile_CI$q25_upper, 2)), ")"),
  `Q75 (95% CI)` = paste0(
    ifelse(is.na(survival_quantile_CI$q75), "NR", round(survival_quantile_CI$q75, 2)),
    " (", ifelse(is.na(survival_quantile_CI$q75_lower), "NR", round(survival_quantile_CI$q75_lower, 2)),
    "&ndash;", ifelse(is.na(survival_quantile_CI$q75_upper), "NR", round(survival_quantile_CI$q75_upper, 2)), ")"),
  check.names = FALSE
)

knitr::kable(qtbl, format = "html", escape = FALSE,
             caption = "Survival Quantiles with 95% Confidence Intervals") %>%
  kableExtra::kable_styling(full_width = FALSE)
```

*NR = Not Reached. The survival estimate did not decline to the specified percentile during the follow-up period. NR in confidence interval bounds indicates the interval could not be estimated &mdash; too few events were observed for the survival curve to reach that quantile threshold within the follow-up period.*

---

## Mean Survival Table

```{r mean-table, echo=FALSE}
mtbl <- data.frame(
  Group           = gsub("group=", "", as.character(survival_mean_CI$group)),
  `Mean Survival` = round(survival_mean_CI$rmean, 2),
  SE              = round(survival_mean_CI$se, 2),
  `95% CI Lower`  = round(survival_mean_CI$lower_ci, 2),
  `95% CI Upper`  = round(survival_mean_CI$upper_ci, 2),
  check.names = FALSE
)

knitr::kable(mtbl, format = "html", escape = FALSE,
             caption = "Restricted Mean Survival Time with 95% Confidence Intervals") %>%
  kableExtra::kable_styling(full_width = FALSE)
```

*Mean survival is bounded by the largest observed event time in each group (restricted mean survival time).*

---

## Interaction Summary

```{r interaction-summary, echo=FALSE}
dat_summary <- data.frame(group_1, group_2, event, time,
                          strata = interaction(group_1, group_2))
int_tbl <- dat_summary %>%
  dplyr::group_by(strata) %>%
  dplyr::summarise(n = dplyr::n(), Events = sum(event), .groups = "drop")

knitr::kable(int_tbl, format = "html", escape = FALSE,
             caption = "Interaction Strata Summary (agecat &times; gender)") %>%
  kableExtra::kable_styling(full_width = FALSE)
```

```{r sparse-warning, echo=FALSE, results='asis'}
sparse <- int_tbl %>% filter(Events < 5)
if (nrow(sparse) > 0) {
  for (s in sparse$strata) {
    cat(paste0("**Warning:** Stratum **", s, "** has fewer than 5 events &mdash; results for this stratum should be interpreted cautiously.\n\n"))
  }
}
```

---

## Pairwise Table (Bonferroni-adjusted)

```{r pairwise-table, echo=FALSE}
dat_pw <- data.frame(group = interaction(group_1, group_2), event = event, time = time)
fit_pairwise_obj <- survminer::pairwise_survdiff(
  survival::Surv(time, event) ~ group,
  data = dat_pw,
  p.adjust.method = "bonferroni"
)
pw_mat <- round(fit_pairwise_obj$p.value, 4)
pw_mat[is.na(pw_mat)] <- "&mdash;"

knitr::kable(pw_mat, format = "html", escape = FALSE,
             caption = "Pairwise Log-Rank p-values (Bonferroni-adjusted)") %>%
  kableExtra::kable_styling(full_width = FALSE)
```

---

## Kaplan-Meier Plot

```{r km-plot, echo=FALSE, fig.width=10, fig.height=7}
graph_lrt(group_1, group_2, event, time)
```

---

## Plain-Language Interpretation

This analysis examined whether survival times differed across combinations of age group and gender among 100 patients in the WHAS100 dataset. The log-rank test with Bonferroni correction was statistically significant (&chi;&sup2;(7) = `r round(lr_chisq, 2)`, p = `r round(lr_p, 4)`), meaning the probability of observing differences this large by chance alone &mdash; if survival were truly identical across all groups &mdash; is approximately `r round(lr_p * 100, 1)`%. We therefore conclude that at least one group's survival experience differs meaningfully from the others.

In practical terms, a clear age gradient is visible: younger patients (ages 32&ndash;70) survived substantially longer on average (restricted mean survival: ~1,579&ndash;2,309 days) compared to the oldest patients (ages 80&ndash;92, mean: ~1,234&ndash;1,262 days), regardless of gender. After Bonferroni correction, only the contrast between [80,92] Male and [32,60) Female reached statistical significance (adjusted p = 0.048). Statistical significance does not, however, imply clinical significance &mdash; the observed differences should be interpreted alongside clinical context.

We are `r (1 - threshold) * 100`% confident that the true median (or mean) survival time for each group lies within the intervals shown in the tables above. Important caveats include: three strata ([60,70) Female, [60,70) Male, [70,80) Male) had fewer than 5 events, which reduces the reliability of pairwise comparisons for those groups; the wide confidence intervals in male strata with small samples reflect considerable uncertainty in those estimates; and the proportional hazards assumption underlying the log-rank test should be formally verified for a definitive assessment.
