---
title: "Claude Statistical Test Evaluation"
output:
  pdf_document:
    toc: true
    toc_depth: '3'
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
header-includes:
  - \usepackage{booktabs}
  - \usepackage{array}
  - \usepackage{ragged2e}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
```

## Prompt Used

> "I have a study with X patients in the treatment arm, X patients in the control arm, X responders in the treatment group, and X responders in the control group. Please use the attached file to run the appropriate statistical test and provide summary statistics."

---

## Multiple Codes

### Chat

```{r multiple-codes-chat}
mc_chat <- data.frame(
  Test = c(
    "test(10,NA,1.2,3)\ninvalid",
    "test(50,40,30,10)\nchisq, dependent",
    "test(50,40,26,19)\nchisq, independent",
    "test(20,10,5,2)\nfisher, independent",
    "test(10,15,10,1)\nfisher, dependent"
  ),
  Correctness = c(
    "T (pointed out NA and 1.2 which will cause error, correct decision)",
    "T (chi sq, test statistic=11.03, p < 0.001. Correctly used p-value labels, correct decision)",
    "T (chi sq, test statistic=0.18, p=0.67. Correctly used p-value labels, correct decision)",
    "T (Fisher, OR=1.33 (small rounding deviation), p=1.0. Correctly used p-value labels, correct decision)",
    "T (Fisher, OR not shown, p=3.37e-6. Correctly used p-value labels, correct decision)"
  ),
  Time = c("11 sec", "26 sec", "25 sec", "27 sec", "38 sec"),
  Format = c(
    "Clean, simple",
    "Clean, with table and sections",
    "Clean, with table and sections",
    "Clean, simple, with a few sections",
    "Clean, with sections and table"
  ),
  `Uses Our Code` = c(
    "Y",
    "N (R not available, but Claude followed my logic)",
    "N (R not available, but Claude followed my logic)",
    "N (R not available, but Claude followed my logic)",
    "N (R not available, but Claude followed my logic)"
  ),
  check.names = FALSE
)

kable(mc_chat, align = "l", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"), font_size = 9) %>%
  column_spec(1, bold = TRUE, width = "9em") %>%
  column_spec(2, width = "16em") %>%
  column_spec(3, width = "4em") %>%
  column_spec(4, width = "10em") %>%
  column_spec(5, width = "10em")
```

---

### Code

```{r multiple-codes-code}
mc_code <- data.frame(
  Test = c(
    "test(10,NA,1.2,3)\ninvalid",
    "test(50,40,30,10)\nchisq, dependent",
    "test(50,40,26,19)\nchisq, independent",
    "test(20,10,5,2)\nfisher, independent",
    "test(10,15,10,1)\nfisher, dependent"
  ),
  Correctness = c(
    "T (correctly specified missing and invalid input, give suggestions)",
    "T (correct p value, test statistics, and decision)",
    "T (correct p value, test statistics, and decision, chisq)",
    "T (correct odds ratio, test statistics, and decision, fisher)",
    "T (correct odds ratio, test statistics, and decision, fisher)"
  ),
  Time = c("19 sec", "15 sec", "15 sec", "15 sec", "18 sec"),
  Format = c(
    "Simple format",
    "Simple, with section and table",
    "Simple and brief with table",
    "Short, brief, with table",
    "Short, brief, with summary table"
  ),
  `Uses Our Code` = c(
    "Y (using our logic)",
    "Y", "Y", "Y", "Y"
  ),
  check.names = FALSE
)

kable(mc_code, align = "l", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"), font_size = 9) %>%
  column_spec(1, bold = TRUE, width = "9em") %>%
  column_spec(2, width = "16em") %>%
  column_spec(3, width = "4em") %>%
  column_spec(4, width = "10em") %>%
  column_spec(5, width = "10em")
```

---

## Single Code

*Parameters: n.tx, n.ctrl, resp.tx, resp.ctrl*

### Chat

```{r single-code-chat}
sc_chat <- data.frame(
  Test = c(
    "test(10,NA,1.2,3)\ninvalid",
    "test(50,40,30,10)\nchisq, dependent",
    "test(50,40,26,19)\nchisq, independent",
    "test(20,10,5,2)\nfisher, independent",
    "test(10,15,10,1)\nfisher, dependent"
  ),
  Accuracy = c(
    "T (flags NA and 1.2 errors and asks for correct inputs)",
    "F (incorrect chisq statistic and p-value, although correct significant conclusion)",
    "T",
    "F (uses Fisher's and gets correct p-value but incorrect OR)",
    "T"
  ),
  Time = c("20 sec", "39 sec", "43 sec", "1 min 2 sec", "46 sec"),
  `Results Format` = c(
    "Brief, clear explanation of errors and offers correction guidelines",
    "Test selection, contingency table, exact hypothesis test output as coded, summary stats, RR/OR interpretation",
    "Study inputs, test selection, contingency table, expected counts, test results, exact conclusion as coded, in plain terms",
    "Study inputs, test selection, test output, summary statistics, conclusion (not as coded), interpretation",
    "Study inputs, test selection, summary statistics, conclusion (not as coded), note (small sample size)"
  ),
  `Uses Our Code` = c(
    "Y (to acknowledge errors via our standards)",
    "N (R not available, manually replicated in Python)",
    "N (R not available, manually executed)",
    "N (R not available, manually executed)",
    "N (R not available, manually executed)"
  ),
  check.names = FALSE
)

kable(sc_chat, align = "l", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"), font_size = 9) %>%
  column_spec(1, bold = TRUE, width = "9em") %>%
  column_spec(2, width = "12em") %>%
  column_spec(3, width = "5em") %>%
  column_spec(4, width = "16em") %>%
  column_spec(5, width = "10em")
```

---

### Code

```{r single-code-code}
sc_code <- data.frame(
  Test = c(
    "test(10,NA,1.2,3)\ninvalid",
    "test(50,40,30,10)\nchisq, dependent",
    "test(50,40,26,19)\nchisq, independent",
    "test(20,10,5,2)\nfisher, independent",
    "test(10,15,10,1)\nfisher, dependent"
  ),
  Accuracy = c("T", "T", "T", "T", "T"),
  Time = c("10 sec", "16 sec", "31 sec", "33 sec", "16 sec"),
  `Results Format` = c(
    "Detects errors based on code and returns error message",
    "Contingency table, statistical test, summary statistics, conclusion (not as coded)",
    "Contingency table, statistical test, summary statistics, conclusion (not as coded)",
    "Contingency table, statistical test, summary statistics, conclusion (not as coded)",
    "Contingency table, statistical test, summary statistics, conclusion (not as coded)"
  ),
  `Uses Our Code` = c("Y", "Y", "Y", "Y", "Y"),
  check.names = FALSE
)

kable(sc_code, align = "l", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"), font_size = 9) %>%
  column_spec(1, bold = TRUE, width = "9em") %>%
  column_spec(2, width = "4em") %>%
  column_spec(3, width = "4em") %>%
  column_spec(4, width = "20em") %>%
  column_spec(5, width = "6em")
```

---

## Conclusion

```{r summary-table}
summary_df <- data.frame(
  Condition = c("Multiple codes, Chat", "Multiple codes, Code", "Single code, Chat", "Single code, Code"),
  `Mean Execution Time` = c("25.4 sec", "16.4 sec", "42 sec", "21.2 sec"),
  Correctness = c("5/5", "5/5", "3/5", "5/5"),
  check.names = FALSE
)

kable(summary_df, align = "l", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position"), full_width = FALSE, font_size = 11) %>%
  column_spec(1, bold = TRUE)
```

We can observe clearly that **multiple codes are more efficient than a single code chunk**, and **Code executes more efficiently compared to Chat**. Chat also has a higher mistake rate when using single code. With multiple codes, the summary Claude provides is usually shorter and more brief compared to those generated using single code.

As a result, using multiple codes is more efficient than using single code, and using Code is more efficient and accurate than Chat.
